Management induced changes of soil organic carbon on global croplands

### 2022-08-19 SOCbudgetPaper/SOCbudget.tex
updated paper version (with hand edited tex file for picture placement issue)
update paper files that where submitted to bg

\begin{figure}[h]
\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/LUtransitions} \caption{Scheme of land-use transition representation. Given an initial land-use pattern (as in this example 2 ha land under natural vegetation and 1 ha of cropland), there are separate $SOC$ stocks for natural vegetation and cropland. While in this example we assume $SOC$ under natural vegetation to be in steady state, the cropland $SOC$ stock approaches its steady state without having reached it yet (a). Upon cropland expansion (in this example half of the natural vegetation is cleared to be used as cropland), $SOC$ stocks on cropland increase due to a transfer of land from natural vegetation (b). Explicitly representing newly converted cropland and existing cropland to account for SOC dynamics (c) leads to the same weighted mean value as averaging $SOC$ stocks (d), due to the linearity of Eq. \ref{eq:SOCstate} and cropland-age independent decay rates (see Eq. \ref{eq:decayrates}).}\label{fig:LUtrans}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data. Panel (a) shows global $\Delta SOC$ between historical land use and potential natural vegetation (PNV). The distribution of total global SOC stocks for the first 30 cm on cropland for the year 2010 is depcited in panel (b). Absolute (c) and relative (d) SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains.}\label{fig:SOCmaps}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=6.3cm]{../ResultNotebooks/Output/Images/SpatialDebtFigure} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile. Panel (a) shows global $\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types. Computing the difference between the $\Delta SOC$ estimate for 2010 and for 1975 (b1) depicts areas of SOC depletion (SOC debt increase, red) and SOC accumulation (SOC debt decline, blue). Panel (b2) shows the of land-use change (LUC) induced change of $\Delta SOC$ between 1975 and 2010, whereas panel (b3) depicts the change due to changing agricultural management (MAN).}\label{fig:SOCdebt}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the cropland system for the year 2010 (in $\unit{MtC}$). Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand. After accounting for losses within the cropland system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues. Large parts of C, however, are mineralized on the field before entering the soil. Additionally, C is transferred to and from global cropland soil stock via land-use change between cropland and natural vegetation. Finally, SOC is mineralized and flows back to the atmosphere.}\label{fig:FlowFig}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in $\unit{GtC}$ for different management scenarios. The stylized scenarios deviate from historical agricultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage). The scenario constManagement combines all three modifications. Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a hypothetical natural vegetation state. Panel (b) shows the carbon inflows from crop residue and manure.}\label{fig:SOCscen}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}. IPCC factors vary a bit with different types of management (low, medium, high). The results of the study span over a wide range -- indicated by the span between the 5th and 95th percentile -- but not overlapping with the IPCCs 2019 factors for tropical moist. The median for most regions is considerably lower than the IPCC factors.}\label{fig:SCFglo}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=7cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area. The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \citep{schaphoff_lpjml4_2018-1} and estimates from \citep{sanderman_soil_2017}. We show values of this study for the year 2010 accounting for the historical land-use dynamics as well as for an hypothetical PNV.}\label{fig:SOCglo}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareSoilGrids2_2010_lu} \caption{Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0.}\label{fig:SOCSoilGrids2}
\end{figure}

\begin{figure}[h]
\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/SOCquantilesSoilGrids2} \caption{Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0.}\label{fig:SOCSoilGrids2Q}
\end{figure}

@@

With the reduced-complexity SOC model we are able to account for agricultural management effects on cropland SOC dynamics within optimization-based IAM frameworks. Reduced input data requirements such as accounting for changes in productivity rather than reproducing the processes that lead to such changes in productivity \citep{elliott_management_trends_2018} will help to explore the role of agricultural management in sustainable development pathway analyses \citep{sorgel_sustainable_2021}.
However, we clearly see that increases in agricultural productivity are not sufficient to create positive net SOC sequestration in cropland soils. More management options that explicitly target the sequestration of C in cropland soils need to be considered. Our open-source model can be expanded to account for additional management options for carbon farming, such as cover crops, agroforestry, or biochar applications.

@@

\begin{figure}[h]
\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones.png} 
\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
\label{fig:CLIMzone}
\end{figure}

\begin{sidewaystable*}[htbp]

@@

         &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\

@@

rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\

@@

others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\

@@

         &                 & \multicolumn{6}{l}{
                                           \begin{minipage}[t]{0.25\columnwidth}

@@

                                                                        \multicolumn{5}{l}{
                                                                            \begin{minipage}[t]{0.25\columnwidth}

@@

\end{sidewaystable*}

@@

\begin{figure}[h]
\includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
\caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510). On the other hand, assuming all SOC to be in land-use specific steady-state already in 1901 (Initial-eq) leads to the opposite effect of an already higher $\Delta SOC$ in 1975 and a less steeper increase till 2010. However, the difference between the dynamic historical management assumption (histManagement) compared to constant management assumption from 1975 on (constManagement) are of a similar size for all initialization choices.}
\label{fig:SOCinit}
\end{figure}

\begin{figure}[h]
\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
\caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
\label{fig:SOCreg}
\end{figure}

\begin{figure}[h]
\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
\caption{Correlation between modeled and measured $SOC$ stocks. Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results. Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
\label{fig:SOCpoint}
\end{figure}

 \begin{figure}[h]
\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
\caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
\label{fig:SOCLPJmL4}
\end{figure}
\noappendix

>>>>

\begin{figure}[h]
\includegraphics[width=8cm]{LUtransitions.png} \caption{Scheme of land-use transition representation. Given an initial land-use pattern (as in this example 2 ha land under natural vegetation and 1 ha of cropland), there are separate $SOC$ stocks for natural vegetation and cropland. While in this example we assume $SOC$ under natural vegetation to be in steady state, the cropland $SOC$ stock approaches its steady state without having reached it yet (a). Upon cropland expansion (in this example half of the natural vegetation is cleared to be used as cropland), $SOC$ stocks on cropland increase due to a transfer of land from natural vegetation (b). Explicitly representing newly converted cropland and existing cropland to account for SOC dynamics (c) leads to the same weighted mean value as averaging $SOC$ stocks (d), due to the linearity of Eq. \ref{eq:SOCstate} and cropland-age independent decay rates (see Eq. \ref{eq:decayrates}).}\label{fig:LUtrans}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=18cm]{4PanelFigure.png} \caption{Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data. Panel (a) shows global $\Delta SOC$ between historical land use and potential natural vegetation (PNV). The distribution of total global SOC stocks for the first 30 cm on cropland for the year 2010 is depcited in panel (b). Absolute (c) and relative (d) SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains.}\label{fig:SOCmaps}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=6.3cm]{SpatialDebtFigure.png} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile. Panel (a) shows global $\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types. Computing the difference between the $\Delta SOC$ estimate for 2010 and for 1975 (b1) depicts areas of SOC depletion (SOC debt increase, red) and SOC accumulation (SOC debt decline, blue). Panel (b2) shows the of land-use change (LUC) induced change of $\Delta SOC$ between 1975 and 2010, whereas panel (b3) depicts the change due to changing agricultural management (MAN).}\label{fig:SOCdebt}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=16cm]{CarbonBudget.png} \caption{Global carbon flows within the cropland system for the year 2010 (in $\unit{MtC}$). Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand. After accounting for losses within the cropland system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues. Large parts of C, however, are mineralized on the field before entering the soil. Additionally, C is transferred to and from global cropland soil stock via land-use change between cropland and natural vegetation. Finally, SOC is mineralized and flows back to the atmosphere.}\label{fig:FlowFig}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=14cm]{ScenarioManageHoriz.png} \caption{(a) Global $\Delta SOC$ in $\unit{GtC}$ for different management scenarios. The stylized scenarios deviate from historical agricultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage). The scenario constManagement combines all three modifications. Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a hypothetical natural vegetation state. Panel (b) shows the carbon inflows from crop residue and manure.}\label{fig:SOCscen}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=16cm]{SCFcomparison.png} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}. IPCC factors vary a bit with different types of management (low, medium, high). The results of the study span over a wide range -- indicated by the span between the 5th and 95th percentile -- but not overlapping with the IPCCs 2019 factors for tropical moist. The median for most regions is considerably lower than the IPCC factors.}\label{fig:SCFglo}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=7cm]{gloComparisonFigure.png} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area. The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \citep{schaphoff_lpjml4_2018-1} and estimates from \citep{sanderman_soil_2017}. We show values of this study for the year 2010 accounting for the historical land-use dynamics as well as for an hypothetical PNV.}\label{fig:SOCglo}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=13cm]{scatterCompareSoilGrids2_2010_lu.png} \caption{Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0.}\label{fig:SOCSoilGrids2}
\end{figure}

\begin{figure}[h]
\includegraphics[width=14cm]{SOCquantilesSoilGrids2.png} \caption{Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0.}\label{fig:SOCSoilGrids2Q}
\end{figure}

@@

With the reduced-complexity SOC model we are able to account for agricultural management effects on cropland SOC dynamics within optimization-based IAM frameworks. Reduced input data requirements such as accounting for changes in productivity rather than reproducing the processes that lead to such changes in productivity \citep{elliott_management_trends_2018} will help to explore the role of agricultural management in sustainable development pathway analyses \citep{sorgel_sustainable_2021}.
However, we clearly see that increases in agricultural productivity are not sufficient to create positive net SOC sequestration in cropland soils. More management options that explicitly target the sequestration of C in cropland soils need to be considered. Our open-source model can be expanded to account for additional management options for carbon farming, such as cover crops, agroforestry, or biochar applications.
\newpage

@@

\begin{figure}[!hp]
\includegraphics[width=13cm]{ClimateZones.png} 
\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
\label{fig:CLIMzone}
\end{figure}

\begin{table*}[!htbp]

@@

         &                                 & 
         \multicolumn{3}{l}{\shortstack[l]{Harvested\\Organs}} & 
         \multicolumn{3}{l}{\shortstack[l]{Above-ground\\Residues}} & 
         \multicolumn{2}{l}{\shortstack[l]{Below-ground\\Residues}} & 
         \multicolumn{3}{l}{\shortstack[l]{Allometric\\coefficients}}                \\

@@

rapeseed   & \begin{tabular}{@{}l@{}}Other oil crops\\(incl rapeseed)\end{tabular} & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\

@@

others     & \begin{tabular}{@{}l@{}}Fruits, Vegetables,\\Nuts\end{tabular}          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\

@@

         &                 & \multicolumn{5}{l}{
                                           \begin{minipage}[t]{0.3\columnwidth}

@@

                                                                        \multicolumn{6}{l}{
                                                                            \begin{minipage}[t]{0.3\columnwidth}

@@

\end{table*}

@@

\begin{figure}[!hp]
\includegraphics[width=12cm]{ScenarioInit.png} 
\caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510). On the other hand, assuming all SOC to be in land-use specific steady-state already in 1901 (Initial-eq) leads to the opposite effect of an already higher $\Delta SOC$ in 1975 and a less steeper increase till 2010. However, the difference between the dynamic historical management assumption (histManagement) compared to constant management assumption from 1975 on (constManagement) are of a similar size for all initialization choices.}
\label{fig:SOCinit}
\end{figure}

\begin{figure}[!hp]
\includegraphics[width=13cm]{regComparisonFigure.png} 
\caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
\label{fig:SOCreg}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=10cm]{scatterComparePointData_2010.png} 
\caption{Correlation between modeled and measured $SOC$ stocks. Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results. Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
\label{fig:SOCpoint}
\end{figure}

 \begin{figure}[ht]
\includegraphics[width=13cm]{scatterCompareLPJmL42010_pnv.png}
\caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
\label{fig:SOCLPJmL4}
\end{figure}
\noappendix
